paper1_figures

### 2023-11-16 scoping_review_short.Rmd
2nd version for feedbacks - 16.11

abstract: |
  Commons represent an alternative social organization outside the classical capitalist model, therefore they could be the needed change toward more sustainable societies and economy. However, in order for a common to be successful, its members must cooperate, i.e. incur a personal sacrifice for the collective good. Despite the importance of cooperation, there is a lack of studies that investigated this topic in commons context. On the other hand, cooperation has been extensively studied in social dilemma experiments, where personal and collective interests are at odds. Since the two contexts share some structural characteristics, the results of social dilemma experiments could be used to gain insights on cooperation in commons.
  The goal of the current scoping review is to identify the factors that impact cooperation in social dilemma situations, to gain a better understanding of what could shape cooperation in commons. A key-word search and a subsequent selection lead to the inclusion of 89 papers, who are used to identify 9 factors that can influence cooperation both at the individual and group level: *gender, social status, group identification, values and personality traits, trust, incentives, communication, social norms and anonymity*. These factors are discussed within the existing literature on commons and cooperation. *More specifically, the influences that these elements can have on each-other are highlighted, showing that the totality of factors should be taken into account when investigating cooperation in commons.* Finally, the practical implication of this work for commons' organization, as well as the review's limitations, are pointed out.

@@

Cooperation can be modified by introducing incentive systems, that can be either punishments or rewards. In the literature, punishment has received a greater attention [@Balliet2011a; @VanDijk2015], and its positive effect on cooperation has been widely demonstrated. In fact, when participants can punish other members for not contributing equally, cooperation and the perceived fairness are higher [@Pfattheicher2018; @Nockur2021].

Several factors moderate this main effect. Punishments are more effective if their implementation is (moderately) costly rather than free [@Balliet2011a; @Egas2008], if the members are friends rather than strangers, if they know that they will interact repeatedly [@Balliet2011a], and if only cooperators have the power to punish defectors [@Xiao2016]. Also, a punishment system increase cooperation mostly in groups where all members have the same income [@Kingsley2016], because in heterogeneous groups a form of solidarity between low-status members arises and encourages free-riding [@Chen2022]. Finally, to have a stronger effect of punishment, the link between behaviour and outcome must be clear [@Xiao2016], and the punishment must be highly effective [@Egas2008].

At the same time, punishment must be applied with carefulness. Ideally, groups should not punish in early stages of the interaction, and have a norm of when and how punishment must be applied [@Rockenbach2019]. Moreover, punishment could entails a second-order social dilemma, since cooperators are better-off by non punishing, if all others do [@Milinski2012]. Also, if there are no rules on who can punish, non-cooperators can punish cooperators as a form of retaliation [@Xiao2016; @VanDijk2015]. A democratic punishment system, where a fine is given only if a majority agree, could solve these problems [@Pfattheicher2018].

On the other side, rewarding systems are less investigated. Generally, they have a positive effect on cooperation [@Balliet2011a], but it is not clear whether they are more efficient than punishment [see @Chen2022], less efficient [see @Milinski2012] or equivalent [see @Balliet2011a]. Rewards have a stronger effect if their implementation is costly and during iterated, rather than single, interactions [@Balliet2011a].

In general, even if they enhance cooperation, incentives have negative sides: they reduce participants' intrinsic motivation to cooperate [@VanDijk2015; @Schlosser2018] by framing the situation as an economic, rather than ethical, dilemma [@VanVugt2009]. They also undermine participants' trust in other members' internal motivation to cooperate. Therefore, if they are removed, cooperation drops [@Irwin2014; @Mulder2006].

@@

The aim of the present paper was to synthesize the existing literature about cooperation in social dilemmas, to identify factors that could impact cooperative behaviours in commons. Through a scoping review approach, we identified nine factors, both at the individual and group level. Our results seem to be in line with others findings about cooperation in commons. For instance, factors such as gender, social status, SVO, incentives, and communication were also found by @Kopelman2002 in their review about lab-based commons, and their effects were the same as the ones reported in the current paper.

Moreover, some of the factors can be related to Ostrom's principles to successfully manage a common resource [@Ostrom1990]. For instance, the clearly defined boundaries can strengthen commoners' social identification, since only when an in- and an out-group exist social identity is salient and therefore an identification is possible [@Tajfel1974]. Even if Ostrom developed this principle on a more organizational level, its secondary effect of creating a social identity can also contribute to improve cooperation. Secondly, communication is essential for the commoners to shape social norms [@Bicchieri2002] and to resolve conflicts. Finally, the principles about the monitoring and the graduated sanction systems underline the importance of incentives. The monitoring system refers to social incentives, and is linked to anonymity, since if people know that their behaviour can be observed, they are more willing to conform to social norms because they worry about others' approval [@Nelissen2013; @Rege2004]. On the other hand, a sanction system can also have a positive impact on cooperation [@Balliet2011a], especially if it is graduate, so if sanctions are not too high at the first violation [@Rockenbach2019].[^4]

@@

Ostrom's principles are especially related to factors at the organizational level. This is understandable when considering her economic approach to the study of the commons. However, taking into account the social dynamics and the characteristics of the commoners is also essential to have a better understanding of how these communities work [@Euler2018; @Nightingale2011]. For this reason, this paper made a step further and identified some individual characteristics affecting the efficiency of commons. Although less studied in commons literature, these individual factors are consistent with existing theories.

First, the reported positive effects of trust on cooperation can be explained through the principle of reciprocity, that posits that individuals cooperate more facing cooperative partners [@Fehr2002; @VanLange1999]. Since in real-life situations it is not always possible to know exactly others' contributions, trust (i.e. beliefs about how much others cooperate) becomes essential to decide how to behave [@Vuolevi2012]. Moreover, trust among participants of a common-pool setting is at core of the framework of @Poteete2010, aimed at explaining why commons can be successful.

Secondly, the influence of SVO on cooperation is explained by the difference between pro-social and pro-self in the importance accorded to the relative benefits. More specifically, pro-social value more others' outcome and equality in outcomes than pro-self [@VanLange1999]. Moreover, the difference in weights accorded to the personal and others' outcome seem to be the only dissimilarity between pro-social and pro-self, because they process the cooperative interaction in the same way [@Fiedler2013].

The reviewed results did not provide a clear conclusion about gender effect on cooperation. However, existing theories would predict that women cooperate more than men, because, according to gender stereotypes, women are more communal and caring, while men are more agentic and performance-oriented [@Eagly2012]. In fact, while managing a natural resource, groups partly composed by women display more solidarity and collaboration than all-men groups [@Westermann2005]. However, every experimental paradigm can trigger the participants' gender identity differently [@Sell2009], and for this reason this effect is subjected to several moderators.

Concerning social status, the analyzed studies showed that usually members of low-status groups cooperated more and showed more solidarity with others low-status members than high-status. Although this effect has not been extensively studied, these results seem to be consistent with the existing literature on social status. First, there are structural differences between high- and low-status group's dynamic, and this makes high-status members more individualistic than low-status [@Lorenzi-Cioldi2002]. High-status individuals also display more system justification and more Social Dominance Orientation than low-status [@Caricati2012; @Vargas-Salfate2018]: they see the social system as fair and they have a preference for inequality between social groups. This is negatively correlated with concern about others and altruism [@Pratto1994].

@@

Until now, the identified factors has been discussed separately. However, their impacts are interconnected, as already anticipated in some reviewed results about moderation effects.

For instance, there are gender differences in trust and SVO: trust is a better predictor of cooperation in women than in men [@Irwin2015; @Dorrough2019], and women are more prevalent between pro-social and less between pro-self [@VanLange1999]. Moreover, SVO and trust are linked [@Ackermann2019], and the expectancy about others' cooperation can even be seen as a part of the construct of SVO [@Bogaert2008]. Therefore, pro-self individuals show less cooperation in part because they expect others not to cooperate.

Trust also depends on social status: high-status individuals display more generalized trust (i.e. general beliefs on others' benevolence), while low-status more depersonalized trust (i.e. more trust in other members of their in-group, just because of their membership) [@Navarro-Carrillo2018]. Since high-status individuals perceive the system as fair [@Caricati2012], it is easier for them to believe in others' benevolent intentions. Conversely, belonging to the group is more important for low-status than for high-status [@Lorenzi-Cioldi2002], therefore they see other group members as more trustworthy, even without knowing them personally.

\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_ OK \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

Secondly, communication impacts several other factors. For instance, it plays an essential role in shaping social norms [@Hogg2006; @Bicchieri2002] and it can improve cooperation by signalling a person's intentions and trustworthiness [@Janssen2015]. Moreover, in a study about organizations, communication was a strong predictor of commitment, that is a component of group identification. This link was especially present for vertical communication, i.e. exchanges about organizational strategies between members at different level of the hierarchy [@Postmes2001]. Therefore, having the opportunity to communicate about the strategies that need to be implemented improve one's identification with the group.

Another factor that influences others is group identification. First, as a consequence of the Social Identity Theory, people who strongly identify with their group are more willing to comply to the group's norms and to pursuit the group's goals [@Tajfel1974]. Therefore, high-identified members need less cooperation-enhancing measures. For instance, incentives are more useful to improve cooperation of low-, rather than high-identified, group members [@VanVugt2009], because high-identified members are already motivated to cooperate. Similarly, given their stronger willingness to act for the group's good, high-identified individuals cooperate even in an anonymous condition, while low-identified acted more for the group goal if their actions were accountable [@Barreto2000].

As already mentioned in the results section, anonymity underlies several other factors. First, when actions are accountable, people are more concerned about others' approval and therefore show more conformity to social norms [@Nelissen2013; @Rege2004]. Also, in large groups [where, as a consequence of the group size, levels of anonymity are higher, @Romano2016] it is more difficult to evaluate the trustworthiness of other participants than in small ones [@Janssen2015].

Finally, economic incentives can have a negative impact on people's internal motivation to cooperate, because they change the perception of the context from ethical to economic [@Janssen2015; @VanVugt2009]. Therefore, even if incentives such as punishments or rewards can be useful to improve cooperation, they reduce the impact of the other factors. For instance, when incentives are present, people think that others cooperate because of them, and this leads to reduced levels of trust [@Irwin2014].

incentives =\> social rewards work better when identifying in the group that can dis-/approve the behaviour @Kopelman2002

##

-   theoretical implications?

Moreover, the analyzed papers showed that people often cooperate. This contradicts the Rational Choice Theory (RCT), who posits that individuals are too self interested to sacrifice their personal benefit to the greater good, unless if forced to do so. Our results are consistent with the existing literature, since RCT has already been refuted by experimental studies, who demonstrated that cooperation was higher than predicted in social dilemmas, and that it was impacted by variables that should not play a role according to RCT [@Ostrom1998].

This review also confirmed the lack of quantitative studies about cooperation in real-life commons: of the 89 analyzed papers, just three had these characteristics, and they all were correlative. Therefore, they could not draw a causal link between the investigated factors and cooperative behaviours.

=\> say that this is a confirmation of a gaps in the literature that should be filled with future researches

-   practical implications: This work will therefore lay the foundations to implement experimental and quantitative studies in real-life commons, which, in turn, will allow to gain insights about cooperation in this context =\> rephrase!

-   In summary, this review has practical implications and can be used as a guide for commons members to design campaigns or activities to improve cooperation. In fact, some of the identified factors can be manipulated in order to create better conditions for cooperation to arise. Moreover, since the different elements are linked, they can impact cooperation both directly and indirectly, via mediation and moderation effects. Based on the consensus of the results and on the quantity of connections with other elements, the factors that seem to have a stronger effect are trust, communication, group identification and accountability of actions. Therefore, interventions based on this elements can be very powerful tools to improve cooperation in commons. On the other hand, some individual factor cannot be manipulated, but must be taken into account when designing an intervention. For instance, commons' composition in terms of gender and social status can have a large influence on the overall cooperation.

# Limitations, future perspectives and conclusion

To conclude, this review summarized existing researches about cooperation in social dilemmas, and therefore provided a set of factors that can help better understanding cooperative behaviours in commons. Moreover, it also highlighted a lack of quantitative studies of this topic in real-life commons. However, this scoping review has also several limitations.

@@

Moreover, although the aim was to use results of social dilemma researches to gain a better understanding of commons, most of the included studies had a laboratory setting. Such a simplification of real-life situations do not allow to exactly reproduce the dynamics present in a commons [@Henry2011; @Cole2010]. In fact, some elements such as cultural factors and specific ecological conditions affecting the decision making process cannot be reproduced in lab experiments [@Janssen2015]. Therefore, this paper offers a good basis about elements influencing cooperation in commons, but future studies should confirm these results in real-life settings. Moreover, this review is not informative about the relative importance of each factor. To shed light on this topic, quantitative studies are needed.

@@

In conclusion, a successful commons relies on the cooperation of its members, but this topic has not yet been extensively studied. Thus, further researches are needed to address the limitations of the current study and to reach a deeper understanding of this topic.

>>>>

abstract: |
  Commons represent an alternative to the classical capitalist model, therefore they could be the needed change toward a more sustainable society and economy. However, in order for a commons to be successful, its members must cooperate, i.e. incur a personal sacrifice for the collective good. Despite the importance of cooperation, there is a lack of studies that investigated this topic in real-life commons. On the other hand, cooperation has been extensively studied in social dilemma experiments, where personal and collective interests are at odds. Since the two contexts share some structural characteristics, the results of social dilemma experiments could be used to gain insights on cooperation in commons.
  The goal of the current scoping review is to identify the factors that impact cooperation in social dilemma situations, in order to better understand what could shape cooperation in commons. A key-word search and a subsequent selection lead to the inclusion of 89 papers, who are used to identify 9 factors that can influence cooperation both at the individual and group level: gender, social status, group identification, values and personality traits, trust, incentives, communication, social norms and anonymity. These factors are discussed within the existing literature on commons and cooperation. More specifically, the influences that these elements can have on each-other are highlighted, showing that the totality of factors should be taken into account when investigating cooperation in commons. Finally, the practical implication of this work for commons' organization, as well as the review's limitations, are pointed out.

@@

Cooperation can be modified by introducing incentive systems, that can be either punishments or rewards. Both have a positive effect on cooperation [@Balliet2011a; ]. However, there are contradictory results about which system is more efficient [see @Chen2022; @Milinski2012].

Several factors moderate this main effect. A meta-analysis found that both rewards and punishments are more effective if their implementation is costly rather than free, and during iterated, rather than single, interactions [@Balliet2011a]. Moreover, punishments work better in groups where all members have the same income [@Kingsley2016], because in heterogeneous groups a form of solidarity between low-status members arises and encourages free-riding [@Chen2022]. Punishments are also more effective if participants are friends [@Balliet2011a] and when only cooperators have the power to punish defectors [@Xiao2016]. Finally, the link between behaviour and outcome must be clear [@Xiao2016], and the punishment must have a significant consequence [@Egas2008].

In general, there are more studies about punishments than about rewards [@VanDijk2015]. When a punishment system is present, participants' perceived fairness is higher [@Nockur2021]. However, punishments must be applied with carefulness. Ideally, groups should not punish in early stages of the interaction, and have a norm of when and how punish [@Rockenbach2019]. Moreover, punishment could entails a second-order social dilemma, since cooperators are better-off by non punishing, if all others do [@Milinski2012]. Also, if there are no rules on who can punish, non-cooperators can punish cooperators as a form of retaliation [@Xiao2016; @VanDijk2015]. A democratic punishment system, where a fine is given only if a majority agree, could solve these problems [@Pfattheicher2018].

Even if they enhance cooperation, incentives have negative sides: they reduce participants' intrinsic motivation [@VanDijk2015; @Schlosser2018] by framing the situation as an economic, rather than ethical, dilemma [@VanVugt2009]. They also undermine participants' trust in other members' internal motivation to cooperate. Therefore, if they are removed, cooperation drops [@Irwin2014; @Mulder2006].

@@

The aim of the present paper was to synthesize the existing literature about cooperation in social dilemmas, in order to identify factors that could impact cooperative behaviours in commons. Through a scoping review approach, we identified nine factors, both at the individual and group level. Our results seem to be in line with others findings about cooperation in commons. For instance, factors such as gender, social status, SVO, incentives, and communication were also found by @Kopelman2002 in their review about lab-based commons, and their effects were the same as the ones reported in the current paper.

Moreover, some of the factors can be related to Ostrom's principles to successfully manage a common resource [@Ostrom1990]. For instance, the clearly defined boundaries can strengthen commoners' social identification, since only when an in- and an out-group exist social identity is salient and therefore an identification is possible [@Tajfel1974]. Even if Ostrom developed this principle on a more organizational level, its secondary effect of creating a social identity can also contribute to improve cooperation. Secondly, communication is essential for the principle about the creation and modification of norms [@Bicchieri2002], and for the one about the arenas to solve conflicts. Finally, the principles concerning the monitoring and the graduated sanction systems underline the importance of incentives. The monitoring system refers to social incentives, and is linked to anonymity, since if people know that their behaviour can be observed, they are more willing to conform to social norms because they worry about others' approval [@Nelissen2013; @Rege2004]. On the other hand, a sanction system can also have a positive impact on cooperation [@Balliet2011a], especially if it is graduate, so if sanctions are not too high at the first violation [@Rockenbach2019].[^4]

@@

Ostrom's principles are especially related to factors at the organizational level. This is understandable when considering her economic approach to the study of the commons. However, taking into account the social dynamics and the characteristics of the commoners is also essential to have a better understanding of how these communities work [@Euler2018; @Nightingale2011]. For this reason, this paper went a step further and identified five individual characteristics affecting the efficiency of commons. Although less studied in commons literature, these individual factors are consistent with existing theories.

First, the reported positive effects of trust on cooperation is explained through the principle of reciprocity, that posits that individuals cooperate more facing cooperative partners [@Fehr2002; @VanLange1999]. Since in real-life situations it is not always possible to know exactly others' willingness to contribute, trust (i.e. beliefs about how much others cooperate) becomes essential to decide how to behave [@Vuolevi2012]. Moreover, trust among participants of a common-pool setting is at core of the framework of @Poteete2010, aimed at explaining why commons can be successful.

Secondly, the influence of SVO on cooperation is due to the difference between pro-social and pro-self in the importance accorded to the relative benefits. More specifically, pro-social value more others' outcome and equality in outcomes than pro-self [@VanLange1999]. Moreover, the difference in weights accorded to the personal and others' outcome seem to be the only dissimilarity between pro-social and pro-self, because they process the cooperative interaction in the same way [@Fiedler2013].

The reviewed results did not provide a clear conclusion about gender effect on cooperation. However, existing theories would predict that women cooperate more than men, because, according to gender stereotypes, women are more communal and caring, while men are more agentic and performance-oriented [@Eagly2012]. In fact, groups partly composed by women display more solidarity and collaboration than all-men groups [@Westermann2005]. However, every experimental paradigm can trigger the participants' gender identity differently [@Sell2009], and for this reason this effect is subjected to several moderators.

Concerning social status, the analyzed studies showed that usually members of low-status groups cooperated more and showed more solidarity with others low-status members than high-status. Although this effect has not been extensively studied, these results seem to be consistent with the existing literature on social status. First, there are structural differences between high- and low-status group's dynamic, that make high-status members more individualistic than low-status [@Lorenzi-Cioldi2002]. High-status individuals also display more system justification and more Social Dominance Orientation than low-status [@Caricati2012; @Vargas-Salfate2018]: they see the social system as fair and they have a preference for inequality between social groups. This is negatively correlated with concern about others and altruism [@Pratto1994].

@@

Until now, the identified factors has been discussed separately. However, their impacts are interconnected, as already anticipated in some reviewed results about moderation effects. For instance, there are gender differences in trust and SVO: trust is a better predictor of cooperation in women than in men [@Irwin2015; @Dorrough2019], and women are more prevalent between pro-social and less between pro-self [@VanLange1999]. Moreover, SVO and trust are linked [@Ackermann2019], and the expectancy about others' cooperation can even be seen as a part of the construct of SVO [@Bogaert2008]. Therefore, pro-self individuals show less cooperation in part because they expect others not to cooperate.

Trust also depends on social status: high-status individuals display more generalized trust (i.e. general beliefs on others' benevolence), while low-status more depersonalized trust (i.e. more trust in other members of their in-group, just because of their membership) [@Navarro-Carrillo2018]. Since high-status individuals perceive the system as fair [@Caricati2012], it is easier for them to believe in others' benevolent intentions. Conversely, belonging to the group is more important for low-status than for high-status [@Lorenzi-Cioldi2002], therefore they see other group's members as more trustworthy, even without knowing them personally.

Secondly, communication impacts several other factors. For instance, it plays an essential role in shaping social norms [@Bicchieri2002] and it can improve cooperation by signalling a person's intentions and trustworthiness [@Janssen2015]. Moreover, the opportunity to exchange about the strategies that need to be implemented improve people's commitment within a group, that is a component of group identification [@Postmes2001]. At the same time, group identification also moderates other effects. First, as a consequence of the Social Identity Theory, high-identified individuals are more willing to comply to the group's norms and to pursue the group's goal [@Tajfel1974]. Therefore, they need less cooperation-enhancing measures such as incentives [@VanVugt2009] or non-anonymous conditions [@Barreto2000]. Moreover, social rewards work better with people who strongly identify with the group that can approve their behaviour [@Kopelman2002].

Finally, as mentioned in the results, anonymity underlies several other factors. First, people comply more to social norms when actions are accountable, because their concern about others' approval [@Nelissen2013; @Rege2004]. Also, in large groups [i.e. in more anonymous situations, @Romano2016] it is more difficult to asses the trustworthiness of other participants than in small ones [@Janssen2015].

## Theoretical implications

The analyzed papers showed that people often cooperate. This contradicts the Rational Choice Theory (RCT), who posits that individuals are too self interested to sacrifice their personal benefit to the greater good, unless if forced to do so. Our general results are therefore consistent with the existing literature, since RCT has already been refuted by experimental studies, who demonstrated that cooperation was higher than predicted in social dilemmas, and that it was impacted by variables that should not play a role according to this theory [@Ostrom1998].

Moreover, this review also corroborated the lack of quantitative studies about cooperation in real-life commons: of the 89 analyzed papers, just three had these characteristics, and they all were correlative. Therefore, they could not draw a causal link between the investigated factors and cooperative behaviours. This confirms a gap in the literature, that could be filled by futures experimental studies about cooperation in commons. Moreover, this review laid the foundations to implement such experiment in real-life commons, by providing a set of factors that could be tested.

## Practical implications

This review can be used as a guide for commons members to design campaigns or activities to improve cooperation. In fact, some of the identified factors can be manipulated in order to create better conditions for cooperation to arise. Moreover, since the different elements are linked, they can impact cooperation both directly and indirectly, via mediation or moderation effects. Based on the quantity of connections with other elements, the factors that seem to have a stronger effect are trust, communication, group identification and accountability of actions. Therefore, interventions based on this elements can be very powerful tools to improve cooperation in commons. On the other hand, some individual factor cannot be manipulated, but must be taken into account when designing an intervention. For instance, commons' composition in terms of gender and social status can have a large influence on the overall cooperation.

# Limitations and future perspectives

To conclude, this review summarized existing researches about cooperation in social dilemmas, and therefore provided a set of factors that can help better understanding cooperative behaviours in commons. It also highlighted a lack of quantitative studies about this topic in real-life commons. It is, however, not without limitations.

@@

Moreover, our aim was to use results of social dilemma researches to gain a better understanding of commons, therefore most of the included studies had a laboratory setting. In fact, this kind of setting is very useful to study social dilemmas because it provides a standard paradigm [@Balliet2011]. However, this does not allow to exactly reproduce the dynamics of a commons [@Henry2011; @Cole2010]: some elements such as cultural factors and specific ecological conditions affecting the decision-making process cannot be reproduced in lab experiments [@Janssen2015]. Therefore, this paper offers a good basis about elements influencing cooperation in commons, but future studies should confirm these results in real-life settings. Moreover, this review is not informative about the relative importance of each factor. To shed light on this topic, quantitative studies are needed.

@@

In conclusion, a successful commons relies on the cooperation of its members, but this topic has not yet been extensively studied. Thus, further researches are needed to address the limitations of the current review and to reach a deeper understanding of this topic.
